---
title: "20190327 Factor Analysis of ASD patients"
author: 'Zanis Fang, UID: ZF2213'
date: "3/27/2019"
output: html_document
---	

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE,
											message = FALSE, warning = FALSE)
library(tidyverse)
```


```{r data_cleaning}
individuals <- read_csv("./individuals.csv")
basic_screen <- read_csv("./basic_medical_screening.csv")

n_unique = c()

for (i in 1:ncol(basic_screen)) {
	n_unique <- append(n_unique, length(unique(basic_screen[[i]])))
}
names(n_unique) <- colnames(basic_screen)
# n_unique[which(n_unique != 2)]

bi_bs <- basic_screen[, which(n_unique %in% c(2, 3, 934, 90549))]
# basic_screen %>% skimr::skim()
# bi_bs %>% skimr::skim()

for (i in 1:ncol(bi_bs)) {
	# structural imputation
	if (length(unique(bi_bs[[i]])) == 2) {
		bi_bs[[i]][is.na(bi_bs[[i]])] <- 0
	}
	# making factor variables
	if (length(unique(bi_bs[[i]])) <= 3) {
		bi_bs[[i]] <- as.factor(bi_bs[[i]])
	}
}

# after_unique <- c()
# for (i in 1:ncol(bi_bs)) {
# 	after_unique <- append(after_unique, length(unique(bi_bs[[i]])))
# }
# after_unique
```


```{r}
sample_id <- bi_bs$subject_sp_id
bi_bs <- bi_bs %>% filter(asd == 1) %>% select(-asd, -subject_sp_id)

imputed_basic <- missMDA::imputeFAMD(bi_bs, ncp = 10)
famd_asd_bs <- FactoMineR::FAMD(imputed_basic$completeObs,
																ncp = 10, graph = FALSE)

```

### Eigenvalues for factor analysis

The eigenvalue of 1st dimension reduced from 8.43 (explain %6.86 variance) to 6.28 (explain %5.12 variance).

2nd dimension increases its eigenvalues and variance explained. 3rd and 4th dimension remains the same.

```{r}
famd_asd_bs$eig %>% as_tibble() %>% mutate(row_num = row_number()) %>%
	ggplot(aes(x = row_num, y = eigenvalue)) +
	geom_point() +
	theme_bw() +
	labs(
		x = "Dimensions",
		y = "Eigenvalues"
	)

asd_ind <- famd_asd_bs$ind$coord %>% as_tibble()


# for (i in 1:ncol(asd_ind)) {
# 	asd_ind[[i]] <- asd_ind[[i]] * famd_asd_bs$eig[i]
# }


```


### Intepretation of dimensions

The top four dimensions can still be intepreted as before, with changes in top variable loadings on **1st** and **4th** dimension.

```{r}
famd_dim1 <- famd_asd_bs$var$contrib %>% as_tibble() %>%
	mutate(variable = rownames(famd_asd_bs$var$contrib)) %>% 
	janitor::clean_names() %>%
	arrange(desc(dim_1)) %>%
	mutate(num_row = row_number())

# famd_dim1 %>%
# 	ggplot(aes(x = num_row, y = dim_1)) +
# 	geom_point() +
# 	geom_vline(xintercept = 17.5) +
# 	labs(title = "Variable Loadings on 1st Dimension",
# 			 x = "Rank",
# 			 y = "Loadings on 1st Dimension") +
# 	theme_bw()

famd_dim1 %>% 
	select(variable, dim_1) %>% top_n(10) %>% knitr::kable()

famd_dim2 <- famd_asd_bs$var$contrib %>% as_tibble() %>%
	mutate(variable = rownames(famd_asd_bs$var$contrib)) %>% 
	janitor::clean_names() %>%
	arrange(desc(dim_2)) %>%
	mutate(num_row = row_number())

# famd_dim2 %>% 
# 	ggplot(aes(x = num_row, y = dim_2)) +
# 	geom_point() + 
# 	geom_vline(xintercept = 9.5) +
# 	labs(title = "Variable Loadings on 2nd Dimension",
# 			 x = "Rank",
# 			 y = "Loadings on 2nd Dimension") +
# 	theme_bw()

famd_dim2 %>% 
	select(variable, dim_2) %>% top_n(10) %>% knitr::kable()

famd_dim3 <- famd_asd_bs$var$contrib %>% as_tibble() %>%
	mutate(variable = rownames(famd_asd_bs$var$contrib)) %>% 
	janitor::clean_names() %>%
	arrange(desc(dim_3)) %>%
	mutate(num_row = row_number())

# famd_dim3 %>% 
# 	ggplot(aes(x = num_row, y = dim_3)) +
# 	geom_point() + 
# 	geom_vline(xintercept = 11.5) +
# 	labs(title = "Variable Loadings on 3rd Dimension",
# 			 x = "Rank",
# 			 y = "Loadings on 3rd Dimension") +
# 	theme_bw()

famd_dim3 %>% 
	select(variable, dim_3) %>% top_n(10) %>% knitr::kable()

famd_dim4 <- famd_asd_bs$var$contrib %>% as_tibble() %>%
	mutate(variable = rownames(famd_asd_bs$var$contrib)) %>% 
	janitor::clean_names() %>%
	arrange(desc(dim_4)) %>%
	mutate(num_row = row_number())

# famd_dim4 %>% 
# 	ggplot(aes(x = num_row, y = dim_4)) +
# 	geom_point() + 
# 	geom_vline(xintercept = 15.5) +
# 	labs(title = "Variable Loadings on 4th Dimension",
# 			 x = "Rank",
# 			 y = "Loadings on 4th Dimension") +
# 	theme_bw()

famd_dim4 %>% 
	select(variable, dim_4) %>% top_n(10) %>% knitr::kable()

```

**First dimension** are mostly sleep/eating/toileting problems, and several developmental disorders. The number of developmental disorder related variables are less than previous analysis.

**Second dimension** are mood related disorders such as anxiety, biopolar depression, etc. (Roughly the same as previous analysis)

**Third dimensions** are mostly birth defects. (Roughly the same as previous analysis)

**Fourth dimensions** are genetic testings (They are independent of birth defects/disorders).


### K-means

Four-cluster is still meaningful.

```{r}
# library(FPDclustering)
library(plotly)
# memory exhausted
# asd_bs_clust <- FPDC(asd_ind[, 1:5], k = 3, nf = 2, nu = 2)
# Silh(asd_bs_clust$probability)

asd_bs_clust <- kmeans(asd_ind, 4)

kmeans_asd <- tibble(
	# cluster1 = fpdclust$probability[, 1],
	# cluster2 = fpdclust$probability[, 2],
  # cluster3 = fpdclust$probability[, 3],
	pc1 = asd_ind[[1]],
	pc2 = asd_ind[[2]],
	pc3 = asd_ind[[3]],
	pc4 = asd_ind[[4]],
	pc5 = asd_ind[[5]],
	pc6 = asd_ind[[6]],
	label = as.factor(asd_bs_clust$cluster)
	)
	# mutate(test = if_else(label == 1, cluster1, cluster2)) %>% 
	# mutate(test = if_else(label == 1, cluster1,
	# 											if_else(label == 2, cluster2, cluster3)))



plot_ly(kmeans_asd,
				x = ~pc1, y = ~pc2, z = ~pc3, color = ~label, size = 0.1,
				marker = list(symbol = 'circle',
											sizemode = 'diameter',
											colorscale = "Viridis"), sizes = 1) %>%
	add_markers() %>% 
  layout(title = "Clusters on 3-dimension scatterplot",
  	scene = list(xaxis = list(title = 'Development&ASD',
  																 range = c(-10, 20)),
  										yaxis = list(title = 'Mood&Depression',
  																 range = c(-20, 20)),
  										zaxis = list(title = 'Birth defects',
  																 range = c(-10, 20))))
```

**K-mean results shows that the clusters are more separated on Mood&Depression rather than Development&ASD.**








