---
title: "Autism Exploratory"
author: 'Zanis Fang, UID: ZF2213'
date: "3/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE,
											message = FALSE, warning = FALSE)
library(tidyverse)

```

### Data cleaning

For modeling, exclude **subject_sp_id, respondent_sp_id, family_id, role, age_at_eval_years** (because of redundancy with age_at_eval_months).

Following variables are removed because they are mostly NA's:
**birth_def_oth_calc, gen_dx_calc, prev_study_calc, gest_age**

Self-reported genetic test results were also removed because they can hardly be quantified, and too many levels to be a categorical variable, besides, some have multiple self-reported test results:
**gen_dx1_self_report, gen_dx2_self_report**

Replace NA's with 0 where NA's are intented to be zero according to code book.



```{r data_cleaning}
individuals <- read_csv("./individuals.csv")
basic_screen <- read_csv("./basic_medical_screening.csv")
asd_scq <- read_csv("./scq.csv")
asd_rbs <- read_csv("./rbs-r.csv")
asd_dcdq <- read_csv("./dcdq.csv")

n_unique = c()

for (i in 1:ncol(basic_screen)) {
	n_unique <- append(n_unique, length(unique(basic_screen[[i]])))
}
names(n_unique) <- colnames(basic_screen)
# n_unique[which(n_unique != 2)]

bi_bs <- basic_screen[, which(n_unique %in% c(2, 3, 934))]
# basic_screen %>% skimr::skim()
# bi_bs %>% skimr::skim()

for (i in 1:ncol(bi_bs)) {
	# structural imputation
	if (length(unique(bi_bs[[i]])) == 2) {
		bi_bs[[i]][is.na(bi_bs[[i]])] <- 0
	}
	# making factor variables
	if (length(unique(bi_bs[[i]])) <= 3) {
		bi_bs[[i]] <- as.factor(bi_bs[[i]])
	}
}

# after_unique <- c()
# for (i in 1:ncol(bi_bs)) {
# 	after_unique <- append(after_unique, length(unique(bi_bs[[i]])))
# }
# after_unique
```


### Factor analysis

One analysis with support from professional autism diagnosis (asd variable), one without support from professional autism diagnosis.

First four dimensions outcompete others.

```{r FAMD}
# slow to estimate number of components
# missMDA::estim_ncpMCA(bi_bs, ncp.max = 10, nbsim = 50, method.cv = "Kfold", threshold = 0.001)

# perhaps it is not worth to spend such long time to impute very small fraction of missing values
imputed_basic <- missMDA::imputeFAMD(bi_bs, ncp = 15)
famd_asd_bs <- FactoMineR::FAMD(imputed_basic$completeObs,
																ncp = 15, graph = FALSE)
famd_no_asd <- FactoMineR::FAMD(imputed_basic$completeObs %>%
																	select(-asd), ncp = 15, graph = FALSE)

# write_csv(famd_no_asd$eig %>% as_tibble(),
# 					"./famd/famd_no_asd_eig.csv")
# write_csv(famd_no_asd$ind$coord %>% as_tibble(),
# 					"./famd/famd_no_asd_ind_coord.csv")
# write_csv(famd_no_asd$ind$contrib %>% as_tibble(),
# 					"./famd/famd_no_asd_ind_contrib.csv")
# write_csv(famd_no_asd$var$coord %>% as_tibble(),
# 					"./famd/famd_no_asd_var_coord.csv")
# write_csv(famd_no_asd$var$contrib %>% as_tibble(),
# 					"./famd/famd_no_asd_var_contrib.csv")

# write_csv(famd_asd_bs$eig %>% as_tibble(),
# 					"./famd/famd_asd_bs_eig.csv")
# write_csv(famd_asd_bs$ind$coord %>% as_tibble(),
# 					"./famd/famd_asd_bs_ind_coord.csv")
# write_csv(famd_asd_bs$ind$contrib %>% as_tibble(),
# 					"./famd/famd_asd_bs_ind_contrib.csv")
# write_csv(famd_asd_bs$var$coord %>% as_tibble(),
# 					"./famd/famd_asd_bs_var_coord.csv")
# write_csv(famd_asd_bs$var$contrib %>% as_tibble(),
# 					"./famd/famd_asd_bs_var_contrib.csv")

famd_no_asd_eig <- famd_no_asd$eig
# famd_no_asd_eig <- read_csv("./famd/famd_no_asd_eig.csv")

famd_no_asd_eig %>% as_tibble() %>% mutate(row_num = row_number()) %>%
	ggplot(aes(x = row_num, y = eigenvalue)) +
	geom_point() +
	theme_bw() +
	labs(
		x = "Dimensions",
		y = "Eigenvalues"
	)

# Choose four components

# another way to do it, quite slow
# for (i in 1:ncol(bi_bs)) {
# 	bi_bs[[i]] <- as.numeric(bi_bs[[i]])
# }
# asd_fa_poly <- psych::fa.parallel.poly(bi_bs)

# dimension coordinate for
# asd_ind <- read_csv("./famd/famd_no_asd_ind_coord.csv")
asd_ind <- famd_no_asd$ind$coord %>% as_tibble()

# skewed distribution
# asd_ind %>% ggplot(aes(x = Dim.1)) + geom_histogram()
# asd_ind %>% ggplot(aes(x = log(Dim.1 + 3.5))) +
# 	geom_histogram(bins = 250)


```


### Variables on first four dimensions

**Using the FAMD result without asd diagnosis information**

Get variable names for top 5 vars for top four dimensions.

```{r top_variables}

# famd_no_asd_var_contrib <- read_csv("./famd/famd_no_asd_var_coord.csv")
famd_no_asd_var_contrib <- famd_no_asd$var$contrib

famd_dim1 <- famd_no_asd_var_contrib %>% as_tibble() %>%
	mutate(variable = rownames(famd_no_asd$var$contrib)) %>% 
	janitor::clean_names() %>%
	arrange(desc(dim_1)) %>%
	mutate(num_row = row_number())

# famd_dim1 %>% 
# 	ggplot(aes(x = num_row, y = dim_1)) +
# 	geom_point() + 
# 	geom_vline(xintercept = 17.5) +
# 	labs(title = "Variable Loadings on 1st Dimension",
# 			 x = "Rank",
# 			 y = "Loadings on 1st Dimension") +
# 	theme_bw()

famd_dim1 %>% 
	select(variable, dim_1) %>% top_n(5) %>% knitr::kable()

famd_dim2 <- famd_no_asd_var_contrib %>% as_tibble() %>%
	mutate(variable = rownames(famd_no_asd$var$contrib)) %>% 
	janitor::clean_names() %>%
	arrange(desc(dim_2)) %>%
	mutate(num_row = row_number())

# famd_dim2 %>% 
# 	ggplot(aes(x = num_row, y = dim_2)) +
# 	geom_point() + 
# 	geom_vline(xintercept = 9.5) +
# 	labs(title = "Variable Loadings on 2nd Dimension",
# 			 x = "Rank",
# 			 y = "Loadings on 2nd Dimension") +
# 	theme_bw()

famd_dim2 %>% 
	select(variable, dim_2) %>% top_n(5) %>% knitr::kable()

famd_dim3 <- famd_no_asd_var_contrib %>% as_tibble() %>%
	mutate(variable = rownames(famd_no_asd$var$contrib)) %>% 
	janitor::clean_names() %>%
	arrange(desc(dim_3)) %>%
	mutate(num_row = row_number())

# famd_dim3 %>% 
# 	ggplot(aes(x = num_row, y = dim_3)) +
# 	geom_point() + 
# 	geom_vline(xintercept = 11.5) +
# 	labs(title = "Variable Loadings on 3rd Dimension",
# 			 x = "Rank",
# 			 y = "Loadings on 3rd Dimension") +
# 	theme_bw()

famd_dim3 %>% 
	select(variable, dim_3) %>% top_n(5) %>% knitr::kable()

famd_dim4 <- famd_no_asd_var_contrib %>% as_tibble() %>%
	mutate(variable = rownames(famd_no_asd$var$contrib)) %>% 
	janitor::clean_names() %>%
	arrange(desc(dim_4)) %>%
	mutate(num_row = row_number())

# famd_dim4 %>% 
# 	ggplot(aes(x = num_row, y = dim_4)) +
# 	geom_point() + 
# 	geom_vline(xintercept = 15.5) +
# 	labs(title = "Variable Loadings on 4th Dimension",
# 			 x = "Rank",
# 			 y = "Loadings on 4th Dimension") +
# 	theme_bw()

famd_dim4 %>% 
	select(variable, dim_4) %>% top_n(5) %>% knitr::kable()


```

#### Intepretation

**First dimension** are mostly ASD and developmental disorders, sleep/eating/toileting problems, but not birth defects or due to another medical conditions.

**Second dimension** are mood related disorders such as anxiety, biopolar depression, etc.

**Third dimensions** are mostly birth defects.

**Fourth dimensions** are genetic testings (They are independent of birth defects/disorders).


### Autism score calculated from top variables of 1st dimensions

Output an file for M-plus software to calculate a "score" for autism.

```{r item_response_theory}

# age (months) at evaluation is removed
asd_dim1 <-
	bi_bs[, which(colnames(bi_bs) %in%
									(famd_dim1 %>%
									 	select(variable, dim_1) %>% 
									 	top_n(17))[[1]])] %>%
	select(-age_at_eval_months)

# asd_dim1 %>% write_csv("./dim1_asd.csv", col_names = FALSE)
asd_irt_dim1 <- read_table("irt_dim1.csv", col_names = FALSE)

```


### K-means clustering

Use K-means to cluster on FAMD individuals' coordinates (15 dimensions).

```{r}
# library(FPDclustering)
library(plotly)
# memory exhausted
# asd_bs_clust <- FPDC(asd_ind[, 1:5], k = 3, nf = 2, nu = 2)
# Silh(asd_bs_clust$probability)

asd_bs_clust <- kmeans(asd_ind, 4)

kmeans_asd <- tibble(
	# cluster1 = fpdclust$probability[, 1],
	# cluster2 = fpdclust$probability[, 2],
  # cluster3 = fpdclust$probability[, 3],
	pc1 = asd_ind[[1]],
	pc2 = asd_ind[[2]],
	pc3 = asd_ind[[3]],
	pc4 = asd_ind[[4]],
	pc5 = asd_ind[[5]],
	pc6 = asd_ind[[6]],
	label = as.factor(asd_bs_clust$cluster),
	asd = as.factor(bi_bs$asd),
	irt_score = asd_irt_dim1$X17
	)
	# mutate(test = if_else(label == 1, cluster1, cluster2)) %>% 
	# mutate(test = if_else(label == 1, cluster1,
	# 											if_else(label == 2, cluster2, cluster3)))



plot_ly(kmeans_asd,
				x = ~pc1, y = ~pc2, z = ~pc3, color = ~label, size = 0.1,
				marker = list(symbol = 'circle',
											sizemode = 'diameter',
											colorscale = "Viridis"), sizes = 1) %>%
	add_markers() %>% 
  layout(title = "Clusters on 3-dimension scatterplot",
  	scene = list(xaxis = list(title = 'Development&ASD',
  																 range = c(-10, 20)),
  										yaxis = list(title = 'Mood&Depression',
  																 range = c(-20, 20)),
  										zaxis = list(title = 'Birth defects',
  																 range = c(-10, 20))))
```

#### Intepretation

Four clusters are quite interpretable.

**Cluster 1**: Individuals without autism but has mood disorder (anxiety, bipolar disorder, etc), without severe birth defects.

**Cluster 2**: Individuals without autism or mood disorder, without severe birth defects.

**Cluster 3**: Individuals who have autism, but without severe birth defects.

**Cluster 4**: Individuals with severe birth defects.

### Effectiveness of using first dimension on diagnosing autism

Color individual's autism diagnosis on 3d scatterplot:

```{r}
plot_ly(kmeans_asd,
				x = ~pc1, y = ~pc2, z = ~pc3, color = ~asd, size = 0.1,
				symbol = ~asd, 
				marker = list(symbol = 'circle',
											sizemode = 'diameter',
											colorscale = "Viridis"), sizes = 1) %>%
	add_markers() %>% 
  layout(
  	title = "Autism diagnosis",
  	scene = list(
  										xaxis = list(title = 'Development&ASD',
  																 range = c(-10, 20)),
  										yaxis = list(title = 'Mood&Depression',
  																 range = c(-20, 20)),
  										zaxis = list(title = 'Birth defects',
  																 range = c(-10, 20))))

```

**Relations between four datasets**

Almost all individuals who took social-communication questionnaire, repetitive behavior test, and/or developmental coordination disorder questionnaire have taken basic medical screening.

```{r combine_four_datasets}

fa_irt_corr <- tibble(
	subject_sp_id = basic_screen$subject_sp_id,
	fa_score = log(asd_ind[[1]] + 3.5),
	mood = asd_ind[[2]],
	birth_defect = asd_ind[[3]],
	irt_score = asd_irt_dim1$X17,
	diag_asd = bi_bs$asd
	)

fa_irt_scq <- full_join(x = fa_irt_corr,
												y = asd_scq[, c(1, 54)] %>%
													rename(scq_score = final_score))

fa_irt_scq_rbs <- full_join(x = fa_irt_scq,
														y = asd_rbs[, c(1, 63)] %>% 
															rename(rbs_score = total_final_score))

fa_irt_scq_rbs_dcdq <-
	full_join(x = fa_irt_scq_rbs,
						y = asd_dcdq[, c(1, 29)] %>% 
							rename(dcdq_score = final_score))
```

```{r correlations, eval = FALSE}

fa_irt_scq_rbs_dcdq %>% ggplot(aes(x = fa_score, y = diag_asd)) +
	geom_point(alpha = 0.01, size = 0.01)
	# stat_smooth(method = "glm", method.args = list(family = binomial(link = "logit")), se = FALSE, fullrange = TRUE, color = "red")

fa_irt_scq_rbs_dcdq %>% ggplot(aes(x = irt_score, y = fa_score, color = diag_asd)) + geom_point(alpha = 0.05)

fa_irt_scq_rbs_dcdq %>% ggplot(aes(x = birth_defect, y = irt_score, color = diag_asd)) + geom_point(alpha = 0.05)


# plot_ly(fa_irt_scq_rbs_dcdq,
# 				x = ~fa_score, y = ~scq_score, z = ~rbs_score,
# 				size = 0.5,
# 				marker = list(symbol = 'circle',
# 											sizemode = 'diameter',
# 											colorscale = "Viridis"), sizes = 1) %>%
# 	add_markers() %>% 
#   layout(scene = list(# xaxis = list(range = c(-10, 15)),
#                      # yaxis = list(range = c(10, 80)),
#                      zaxis = list(range = c(0, 120))))


```


**ROC curves**

I use ROC curves to check how comparable the sensitivity and specificity are between **1st dimension**, **top 16 variables from 1st dimension** generated from **Basic Medical Screening** and **Social-Communication Questionnaire**, **Developmental Coordination Disorder** and **Repetitive Behavior Scaled**. However, since all individuals who took social-communication and developmental coordination disorder questionnaires are autism patients, I cannot generate their ROC curves.

Red: IRT (top 16 variables on 1st dimension alone)
Blue: 1st dimension score
Black: Repetitive Behavior Scale

```{r roc_curves}
roc_fa_dim1 <-
	pROC::plot.roc(kmeans_asd$asd, kmeans_asd$pc1,
								 col = "blue", print.thres = TRUE)

roc_irt_dim1 <-
	pROC::plot.roc(kmeans_asd$asd, kmeans_asd$irt_score,
								 add = TRUE, col = "red", print.thres = TRUE)


roc_scq <- pROC::plot.roc(fa_irt_scq_rbs_dcdq$diag_asd,
							 fa_irt_scq_rbs_dcdq$scq_score,
							 add = TRUE, col = "black", print.thres = TRUE)


```

**They have comparable specificity, but much less sensitivity**

### IQ, Epilepsy

```{r}

bi_bs %>%
	mutate(severity = log(asd_ind$Dim.1 + 3.5),
				 asd = basic_screen$asd) %>%
	ggplot(aes(x = severity, y = asd, color = asd)) +
	geom_point(alpha = 0.01, size = 0.01, show.legend = FALSE) +
	theme_bw() +
	theme(panel.grid.major = element_blank(),
				panel.grid.minor = element_blank()) +
	labs(x = "Severity of ASD",
			 y = "ASD")


bi_bs %>%
	mutate(severity = log(asd_ind$Dim.1 + 3.5)) %>%
	ggplot(aes(x = severity, y = dev_id, color = dev_id)) +
	geom_point(alpha = 0.01, size = 0.01, show.legend = FALSE) +
	theme_bw() +
	theme(panel.grid.major = element_blank(),
				panel.grid.minor = element_blank()) +
	labs(x = "Severity of ASD",
			 y = "Intellectual disability")

bi_bs %>%
	mutate(severity = log(asd_ind$Dim.1 + 3.5)) %>%
	ggplot(aes(x = severity, y = gen_test_id, color = gen_test_id)) +
	geom_point(alpha = 0.1, show.legend = FALSE) +
	theme_bw() +
	theme(panel.grid.major = element_blank(),
				panel.grid.minor = element_blank()) +
	labs(x = "Severity of ASD",
			 y = "Genetic test for Intellectual disability")


bi_bs %>%
	mutate(severity = log(asd_ind$Dim.1 + 3.5)) %>%
	ggplot(aes(x = severity, y = neuro_sz, color = neuro_sz)) +
	geom_point(alpha = 0.01, show.legend = FALSE) +
	theme_bw() +
	theme(panel.grid.major = element_blank(),
				panel.grid.minor = element_blank()) +
	labs(x = "Severity of ASD",
			 y = "Seizure or epilepsy")

bi_bs %>% mutate(severity = log(asd_ind$Dim.1 + 3.5)) %>%
	ggplot(aes(x = severity, y = sex, color = sex)) +
	geom_point(alpha = 0.01, show.legend = FALSE) +
	theme_bw() +
	theme(panel.grid.major = element_blank(),
				panel.grid.minor = element_blank()) +
	labs(x = "Severity of ASD",
			 y = "Sex")


```

